---
title: 'Homework 1'
author: "Frank Marino"
date: '2026-01-16'
output:
  html_document:
    theme: readable
---


```{r, echo = FALSE}
# clear environment
rm(list = ls())
```

## Load Libraries
```{r, warning = FALSE, message = FALSE}
# Load any R Packages you may need
require(dplyr)
require(ggplot2)
```

## Exercise 1
The state of Arizona records the number of automobile deaths two years ago in each of the state's 14 counties:

-   30, 15, 183, 9, 7, 39, 28, 30, 76, 13, 21, 18, 323, 3

### (a)

Using R, enter these data into a variable called `autodths`.

```{r}
# Type your code here
autodths <- data.frame(
  county = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14),
  county_death = c(30, 15, 183, 9, 7, 39, 28, 30, 76, 13, 21, 18, 323, 3)
)
```


### (b)

What is the mean number of deaths per county?
```{r}
# Type your code here
mean_death <- mean(autodths$county_death)
```
The mean number of deaths per county is `r round(mean_death, 1)`.


### (c)

What is the median number of deaths per county?
```{r}
# Type your code here
median_death <- median(autodths$county_death)
```
The median number of deaths per county is `r round(median_death, 2)`.




```{r}
colors <- ifelse(autodths$county_death == max(autodths$county_death), 
                 "red",
                 ifelse(autodths$county_death == min(autodths$county_death),
                        "green",
                        "grey"))

chart <- barplot(autodths$county_death,
        names.arg = autodths$county,
        xlab = "County",
        ylab = "Number of Automobile Deaths",
        main = "Number of Automobile Deaths in Arizona Counties (2024)",
        col = colors,
        cex.lab = 1.5
        )
abline(h = mean(autodths$county_death),
                      col = "blue",
                      lty = 2,
                      lwd = 2
                      )
abline(h = median(autodths$county_death),
       col = "red",
       lty = 2,
       lwd = 2
       )
legend("topleft",
       legend = c(paste("Mean =", round(mean_death, 1)),
                  paste("Median =", round(median_death, 2))),
       col = c("blue", "red"),
       lty = c(2, 2),      # Both dashed
       lwd = c(2, 2),      # Both width 2
       bty = "n",          # No box
       cex = 0.9)
```



* * *

## Exercise 2
In a survey of students several years ago, one question asked for each student's political preference (Democrat (D), Independent (I), or Republican (R)). Here are the 14 responses:

-   D, I, I, D, I, D, I, I, I, D, I, R, D, D

### (a)

Using R, enter these data into a variable called `polpref`.

```{r}
# Type your code here
polpref <- data.frame(
  preference = c("D", "I", "I", "D", "I", "D", "I", "I", "I", "D", "I", "R", "D", "D")
)
```

### (b)

Use the `tally()` function to determine the most popular choice of political preference.

```{r}
# Type your code here
polpref <- polpref %>%
  group_by(preference) %>%
  tally()
polpref
```

```{r}
colors <- ifelse(polpref$preference == "D", "blue", ifelse(polpref$preference == "I", "purple", "red"))
barplot(polpref$n,
        names.arg = polpref$preference,
        xlab = "Political Preference",
        ylab = "Number of Students",
        main = "Political Preference Among College Students",
        col = colors)
```


 
> The most popular choice of political preference is Independent. 

* * *

## Exercise 3
### (a)
Create a Canva Presentation with two slides:

- Slide 1: Create a data visualization for the data in Exercise 1
- Slide 2: Create a data visualization for the data in Exercise 2

### (b)
<div style="position: relative; width: 100%; height: 0; padding-top: 56.2500%;
 padding-bottom: 0; box-shadow: 0 2px 8px 0 rgba(63,69,81,0.16); margin-top: 1.6em; margin-bottom: 0.9em; overflow: hidden;
 border-radius: 8px; will-change: transform;">
  <iframe loading="lazy" style="position: absolute; width: 100%; height: 100%; top: 0; left: 0; border: none; padding: 0;margin: 0;"
    src="https://www.canva.com/design/DAG-sVwLV2w/6fhNqd-utjrbaLwMPo9-RA/view?embed" allowfullscreen="allowfullscreen" allow="fullscreen" data-external="1">
  </iframe>
</div>
<a href="https:&#x2F;&#x2F;www.canva.com&#x2F;design&#x2F;DAG-sVwLV2w&#x2F;6fhNqd-utjrbaLwMPo9-RA&#x2F;view?utm_content=DAG-sVwLV2w&amp;utm_campaign=designshare&amp;utm_medium=embeds&amp;utm_source=link" target="_blank" rel="noopener">Automobile Deaths in Arizon Counties (2024)</a> by Frank Marino


Embed your Canva Presentation Here: **INSERT IFRAME EMBED CODE. DON'T FORGET to add in `data-external="1"` at the end of your `<iframe>` or else R-Markdown will not display properly.**

### (c)
Trouble embedding? No worries! Provide your Canva link here as a back-up

URL Link to Canva Site: [Frank Marino - Homework 1](https://www.canva.com/design/DAG-sVwLV2w/6fhNqd-utjrbaLwMPo9-RA/view)

* * *

## All done!

Knit the completed R Markdown file as a HTML document (click the "Knit" button at the top of the script editor window) and upload it to the submission portal on Canvas.
