---
title: | 
 | Data Visualization Project 1
 | Infographic
subtitle: |
 | [ENTER TITLE HERE]
author: | 
 | [Frank Marino]
 | Northeastern University
 | INSH 5302: Information Design and Visual Analytics
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    code_download: true
    fig_caption: yes
    df_print: kable
    theme: lumen
    toc: yes
    toc_float:
      collapsed: false
---

**DELETE ALL TEMPLATE PROMPTS WHEN FINISHED**

# Introduction

> Provide a general question you hope to accomplish with this project.

Provide a brief introduction to the data analysis you conducted, the visualizations you created, and the information you displayed. This section should be between 1 to 2 paragraphs in length, where a paragraph contains at least 4-6 sentences. Delete these instructions when completed.

## Data
### Load Libraries and Packages
*Load any necessary R packages needed for your project. This should include the either the moderndive package or the ISLR package for your data. Delete these instructions when completed.*

```{r}
rm(list = ls())

```

```{r, message=FALSE, warning=FALSE}
# Type your code here...
require(moderndive)
require(dplyr)
require(tidyselect)
library(mosaic)
require(tibble)
require(sqldf)
DD_SB <- DD_vs_SB
```

### Description of data
*Please give a very short description of the data set you chose from the list along with the package this data is found in and why you chose this data set. Delete these instructions when completed.*

> R Package:

> Data set: 

### Load and Clean Data
*Clean your data set to remove null values and to pare down the data set so it only includes the columns (which are your variables) used in your analysis. Delete these instructions when completed.*
```{r, message=FALSE, warning=FALSE}
# Type your code here...
# SQL to combine data??? 
dd_sb_clean <- DD_SB %>%
  filter(
    shops >= 1,
    !is.na(median_income),
    !is.na(shop_type)
  )
dd_sb_clean <- dd_sb_clean %>%
  select(-FIPS)
```

### Preview Data
*Show the first 10 rows of data by using the 'head()' function. Delete these instructions when completed.*
```{r, message=FALSE, warning=FALSE}
# Type your code here...
head(dd_sb_clean,10)
```

### Variables
*Using the names() function, show the names of all the columns (i.e.  variables) in your pared data set and provide a brief explanation of what each variable is. Delete these instructions when completed.*
```{r, message = FALSE, warning = FALSE}
# Type your code here...
names(county_wide)
```

The variables featured in this data visualization project are:

1. County - The county in which the store is located.
2. Median_income - The median household income of the country in which the store is located.
3. Total Population - The population of the county in which the store is located. 
4. Shop_type - Whether the store is a Dunkin Donuts or a Starbucks. 
5. Total Shops - The number of locations for each shop_type in the county.

* * *

# Data Analysis

## Required Analysis

### Chart/Graph 
*Using the appropriate functions, use this space to create the necessary plots for your infographic. Delete these instructions when completed.*
```{r, message = FALSE, warning = FALSE}
library(ggplot2)
ggplot(dd_sb_clean, aes(x = shop_type, y = median_income, fill = shop_type)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.7) +
  geom_jitter(aes(color = shop_type), width = 0.2, alpha = 0.4) +
  scale_fill_manual(
    values = c(
      "starbucks" = "#006241",
      "dunkin_donuts" = "#F5821F"
    )
  ) +
  scale_color_manual(
    values = c(
      "starbucks" = "#006241",
      "dunkin_donuts" = "#F5821F"
    )
  ) +
  scale_x_discrete(labels = c(
    "starbucks" = "Starbucks",
    "dunkin_donuts" = "Dunkin Donuts"
  )) +
  labs(
    title = "Median Income by Store Type",
    x = "Store Type",
    y = "Median Income"
  ) +
  theme_minimal()

```

### Summary Statistics
*Using the favstats() function, calculate the necessary statistics you used to create your infographic. Delete these instructions when completed.*
```{r, message = FALSE, warning = FALSE}
favstats(~median_income, data =dd_sb_clean)
favstats(~shop_type, data = dd_sb_clean)
```

## Additional Analyses
*If you conducted any other analyses of your data set to gather the data in your project, use this section to conduct your research. Delete these instructions when completed.*
```{r, message = FALSE, warning = FALSE}
wilcox.test(median_income ~ shop_type, data = dd_sb_clean)
tapply(dd_sb_clean$median_income, dd_sb_clean$shop_type, median)

```

* * *

# Infographic

*Embed your Canva Presentation Here: **INSERT IFRAME EMBED CODE. DON'T FORGET to add in `data-external="1"` at the end of your `<iframe>` or else R-Markdown will not display properly.***

URL Link to Canva Site: [Your Name - Data Viz Project 1](**COPY AND PASTE YOUR CANVA DESIGN URL HERE**)

* * *

# References
*Include any references that you might have used to construct your graphic. This may include, but is not limited to citations for: graphics used, coding techniques, data collection, R-Markdown/RStudio documentation, etc. Delete this when complete.*

1. Reference 1
2. Reference 2
3. Reference 3
4. Reference 4
5. Reference 5

* * *

# All done!
Knit the completed R Markdown file as a HTML document (click the "Knit" button at the top of the script editor window) and upload it to the submission portal on Canvas.