---
title: 'Visual Project II'
author: "Frank Marino"
date: 'March 1, 2026'
output:
  html_document: 
    code_download: true
    fig_caption: true
    df_print: kable
    theme: lumen
    toc: true
    toc_float: true
    code_folding: null
    toc_depth: 4
---
## Load Libraries
```{r, warning = FALSE, message = FALSE}
# Load any R Packages you may need
rm(list = ls())
library(mosaic)
library(tidyverse)
library(treemap)
library(devtools)
library(dplyr)
library(ggplot2)
library(plotly)
library(readxl)
```


```{r}
df <- read_xlsx("Summer_Olympic_medallists_1896-2008.xlsx")
# Only focusing on sheet 1 -- Using other sheets strictly for checking analysis check. 

# Winter Olympics separated to every 2 years in 1986
# Filter out figure skating, ice hockey,  -- Winter Olympic Games created in 1924. 
```


### Descriptive Statistics:
```{r}
summary(df)
head(df)
str(df)
```

```{r}
colSums(is.na(df))
```


### Filter Winter Sports:
The Winter Olympic Games were officially created in 1924, so prior to this, Winter sports, like Ice Hockey and Figure Skating, were included in the overall Olympic Games. I will be filtering out these two sports to focus solely on sports performed during the Summer Olympic Games.
```{r}
df.summer <- df %>%
  filter(!Sport %in% c("Ice Hockey", "Skating"))
```

### Statistics by Medal:
#### Gold:
```{r}
df.gold <- df.summer %>%
  filter(Medal == "Gold") %>%
  group_by(NOC) %>%
  summarise(Gold_Count = n()) %>%
  arrange(desc(Gold_Count))
```


#### Silver:
```{r}
df.silver <- df.summer %>%
  filter(Medal == "Silver") %>%
  group_by(NOC) %>%
  summarize(Silver_Count = n()) %>%
  arrange(desc(Silver_Count))

```

#### Bronze:
```{r}
df.bronze <- df.summer %>%
  filter(Medal == "Bronze") %>%
  group_by(NOC) %>%
  summarize(Bronze_Count = n()) %>%
  arrange(desc(Bronze_Count))
```


### Statistics by Athlete:
```{r}
df.athlete <- df.summer %>%
  group_by(Athlete, NOC) %>%
  summarize(Athlete_Win = n()) %>%
  arrange(desc(Athlete_Win))
```

```{r}
sum(df.summer$Medal == "Gold")
sum(df.summer$Medal == "Silver")
sum(df.summer$Medal == "Bronze")
```

Total Country Wins
```{r}
df.county <- df.summer %>%
  group_by(NOC) %>%
  summarize(Country_Total = n()) %>%
  arrange(desc(Country_Total))
```



Total Events:
```{r}
nrow(df.summer)
```


### Men vs Women
```{r}
sum(df.summer$Gender == "Men")
sum(df.summer$Gender == "Women")
```

```{r}
df.women1900 <- df.summer %>%
  filter(Gender == "Women") %>%
  filter(Edition == 1900)

nrow(df.women1900)

df.women2008 <- df.summer %>%
  filter(Edition == 2008, Gender == "Women",)

nrow(df.women2008)
```



