---
title: | 
 | Data Visualization Project 2
 | Tableau Story
subtitle: |
 | [Historical Patterns of Summer Olympic Games]
author: | 
 | [Frank Marino]
 | Northeastern University
 | INSH 5302: Information Design and Visual Analytics
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    code_download: true
    fig_caption: yes
    df_print: kable
    theme: lumen
    toc: yes
    toc_float:
      collapsed: false
---

**DELETE ALL TEMPLATE PROMPTS WHEN FINISHED**

# Introduction

> Provide a general question you hope to accomplish with this project.

Provide a brief introduction to the data analysis you conducted, the visualizations you created, and the information you displayed. This section should be between 1 to 2 paragraphs in length, where a paragraph contains at least 4-6 sentences. Delete these instructions when completed.

## Data
### Load Libraries and Packages
*Load any necessary R packages needed for your project. Delete these instructions when completed.*
```{r, message=FALSE, warning=FALSE}
rm(list = ls())
library(mosaic)
library(tidyverse)
library(treemap)
library(devtools)
library(dplyr)
library(ggplot2)
library(plotly)
library(readxl)
```

### Description of data
*Please give a very short description of the data set you chose from the list along with the package this data is found in and why you chose this data set. Delete these instructions when completed.*

> Topic: [Choose the topic from this list that corresponds to your data set: Business, Entertainment, Sports, Public Data, Education, Government, Science, Lifestyle, Technology, OR Health]

> Data set: [Type name of data set here]

### Load and Clean Data
*Clean your data set to remove null values and to pare down the data set so it only includes the columns (which are your variables) used in your analysis. Delete these instructions when completed.*
```{r, message=FALSE, warning=FALSE}
df <- read_xlsx("Summer_Olympic_medallists_1896-2008.xlsx")
# Only focusing on sheet 1 -- Using other sheets strictly for checking analysis check. 

# Winter Olympics separated to every 2 years in 1986
# Filter out figure skating, ice hockey,  -- Winter Olympic Games created in 1924. 
df.summer <- df %>%
  filter(!Sport %in% c("Ice Hockey", "Skating")) %>%
  mutate(NOC = recode(NOC, "URS" = "RUS"))

df.medals <- df.summer %>%
  distinct(Edition, Event, NOC, Medal, Gender, Sport, Discipline, Event_gender)

colSums(is.na(df.summer))

```

### Preview Data
*Show the first 10 rows of data by using the 'head()' function. Delete these instructions when completed.*
```{r, message=FALSE, warning=FALSE}
head(df.summer,10)
```

### Variables
*Using the names() function, show the names of all the columns (i.e.  variables) in your pared data set and provide a brief explanation of what each variable is. Delete these instructions when completed.*
```{r, message = FALSE, warning = FALSE}
names(df.summer)
```

The variables featured in this data visualization project are:

1. City - The city in which the Summer Olympics were held.
2. Edition - The year in which the Summer Olympics were held.
3. Sport - The overall category of sport.
4. Discipline - The specific event of a sport.
5. Athlete - The name of the athlete that won the corresponding medal. 
6. NOC - Abbreviation of participating country. 
7. Gender - The gender of the individual participating in the event.
8. Event - The title of the event in each sport and discipline category.
9. Event_gender - The gender category for the event.
10. Medal - The winning medal awarded to the coinciding Athlete. 

* * *

# Data Analysis

## Required Analysis

### Chart/Graph 
*Using the appropriate functions, use this space to create the necessary plots for your project. Delete these instructions when completed.*
```{r, message = FALSE, warning = FALSE}



```

### Summary Statistics
*Using the favstats() function, calculate the necessary statistics you used to create your project. Delete these instructions when completed.*
```{r, message = FALSE, warning = FALSE}
df.gold <- df.summer %>%
  filter(Medal == "Gold") %>%
  group_by(NOC) %>%
  summarise(Gold_Count = n(), .groups = "drop") %>%
  arrange(desc(Gold_Count))

df.country <- df.medals %>%
  group_by(NOC) %>%
  summarise(Country_Total = n(), .groups = "drop") %>%
  arrange(desc(Country_Total))


df.athlete <- df.summer %>%
  group_by(Athlete, NOC, Gender) %>%
  summarise(Athlete_Win = n(), .groups = "drop") %>%
  arrange(desc(Athlete_Win))

df.gender_edition <- df.medals %>%
  group_by(Edition, Gender) %>%
  summarise(Total = n(), .groups = "drop")

nrow(df.summer)

df.women1900 <- df.medals %>%
  filter(Gender == "Women", Edition == 1900)

nrow(df.women1900)

df.women2008 <- df.medals %>%
  filter(Edition == 2008, Gender == "Women")

nrow(df.women2008)
```




```{r message = FALSE, warning = FALSE}

favstats(df.gold$Gold_Count)
favstats(df.country$Country_Total)
favstats(df.athlete$Athlete_Win)


summary_stats <- df.gold %>%
summarise(
  mean_gold = mean(Gold_Count),
  median_gold = median(Gold_Count),
  sd_gold = sd(Gold_Count),
  min_gold = min(Gold_Count),
  max_gold = max(Gold_Count),
  q1 = quantile(Gold_Count, 0.25),
  q3 = quantile(Gold_Count, 0.75),
  n = n()
)
summary_stats

df.medals %>%
  summarise(num_events = n_distinct(Event))
```
```{r}
ggplot(df.gold %>% top_n(20, Gold_Count),
       aes(x = reorder(NOC, Gold_Count), y = Gold_Count, fill = Gold_Count)) +
  geom_col() +
  coord_flip() +
scale_fill_gradient(low = "#FFD700", high = "#8B6508") +
labs(title = "Top 20 Countries by Gold Medals",
       x = "Country",
       y = "Gold Medals")
```


```{r}
# Total medals per edition
df.medals %>%
  group_by(Edition) %>%
  summarise(Total = n(), .groups = "drop") %>%
  ggplot(aes(x = Edition, y = Total)) +
  geom_line(color = "#EFBF04", linewidth = 1.5) +
  geom_point(color = "#977547", size = 3) +
  geom_vline(xintercept = 1900, linetype = "dashed", color = "gray40") +
  annotate("text", x = 1900, y = 550, label = "Women First Compete",
           angle = 90, vjust = -0.5, color = "deeppink") +
  labs(title = "Expansion of the Summer Olympic Games",
       subtitle = "Total medals awarded per Olympic edition",
       x = "Olympic Year",
       y = "Total Medals") +
  theme_minimal()
```
```{r}
ggplot(df.gender_edition, aes(x = Edition, y = Total, color = Gender)) +
  geom_line(linewidth = 0.75) +
  geom_point() +
  scale_color_manual(values = c("Men" = "blue", "Women" = "deeppink")) +
  geom_vline(xintercept = 1900, linetype = "dashed", color = "gray40") +
  annotate("text", x = 1900, y = 370, label = "Women First Compete",
           angle = 90, vjust = -0.5, color = "deeppink") +
  labs(title = "Olympic Medals by Gender Over Time",
       x = "Olympic Year",
       y = "Total Medals")
```
```{r}
df.country_medals <- df.medals %>%
  filter(NOC %in% c("USA", "RUS", "GBR", "GER", "FRA")) %>%
  group_by(NOC, Medal) %>%
  summarise(Count = n(), .groups = "drop") %>%
  mutate(Medal = factor(Medal, levels = c("Gold", "Silver", "Bronze")))

ggplot(df.country_medals, aes(x = NOC, y = Count, fill = Medal)) +
  geom_col() +
  scale_fill_manual(values = c("Gold" = "#EFBF04",
                               "Silver" = "#758A9c",
                               "Bronze" = "#977547")) +
  labs(title = "Medal Breakdown for Top Performing Countries",
       x = "Country",
       y = "Medal Count")
```



```{r}
library(plotly)

p <- df.medals %>%
  group_by(Edition, Sport) %>%
  summarise(Total = n(), .groups = "drop") %>%
  ggplot(aes(x = Edition, y = Total, color = Sport)) +
  geom_line() +
  labs(title = "Sport Popularity Over Time",
       x = "Edition",
       y = "Total Medals")

graph <- ggplotly(p)
graph

htmlwidgets::saveWidget(graph, "index.html", selfcontained = FALSE)
```

```{r}
head(df.athlete, 10)
```


```{r}
library(tidyverse)
library(gganimate)

# Medal totals by country and edition
df.country_year <- df.medals %>%
  group_by(Edition, NOC) %>%
  summarise(Medals = n(), .groups = "drop")

# Create all country-year combinations so cumulative totals work correctly
all_years <- sort(unique(df.summer$Edition))
all_countries <- sort(unique(df.summer$NOC))

df.country_year_full <- expand.grid(
  Edition = all_years,
  NOC = all_countries
) %>%
  left_join(df.country_year, by = c("Edition", "NOC")) %>%
  mutate(Medals = replace_na(Medals, 0)) %>%
  arrange(NOC, Edition) %>%
  group_by(NOC) %>%
  mutate(Cumulative_Medals = cumsum(Medals)) %>%
  ungroup()

# Keep only top 10 countries in each year
df.top10 <- df.country_year_full %>%
  group_by(Edition) %>%
  slice_max(order_by = Cumulative_Medals, n = 10, with_ties = FALSE) %>%
  ungroup()

# Make country order change within each frame
df.top10 <- df.top10 %>%
  group_by(Edition) %>%
  arrange(Cumulative_Medals) %>%
  mutate(rank = row_number()) %>%
  ungroup()


df.women_year <- df.medals %>%
  filter(Gender == "Women") %>%
  group_by(Edition) %>%
  summarise(Female_Medals = n(), .groups = "drop")
model <- lm(Female_Medals ~ Edition, data = df.women_year)
summary(model)
coef(model)["Edition"]


df.women_participation <- df.summer %>%
  filter(Gender == "Women") %>%
  group_by(Edition) %>%
  summarise(Female_Athletes = n_distinct(Athlete), .groups = "drop")

model_participation <- lm(Female_Athletes ~ Edition, data = df.women_participation)

round(coef(model_participation)["Edition"], 2)

library(grid)

value_1900 <- df.women_participation %>%
  filter(Edition == 1900) %>%
  pull(Female_Athletes)

value_2008 <- df.women_participation %>%
  filter(Edition == 2008) %>%
  pull(Female_Athletes)

ggplot(df.women_participation, aes(x = Edition, y = Female_Athletes)) +
  geom_line() +
  geom_smooth(method = "lm", se = FALSE, color = "deeppink") +
  geom_vline(xintercept = 1900, linetype = "dashed", color = "gray40") +
  annotate("text", x = 1900, y = 370, label = "Women First Compete",
           angle = 90, vjust = -0.5, color = "deeppink") +

  geom_segment(aes(x = 2006, xend = 2007.5,
                   y = value_2008 + 100, yend = value_2008+ 15),
               arrow = arrow(length = unit(0.3, "cm")),
               color = "deeppink") +

  annotate("text", x = 2000, y = value_2008 + 150,
           label = paste(value_2008, "Athletes"),
           color = "deeppink") +

  geom_segment(aes(x = 1904, xend = 1900.5,
                   y = value_1900 + 110, yend = value_1900 + 15),
               arrow = arrow(length = unit(0.3, "cm")),
               color = "deeppink") +

  annotate("text", x = 1908, y = value_1900 + 150,
           label = paste(value_1900, "Athletes"),
           color = "deeppink") +

  labs(title = "Women Participation Rates Across Olympic History",
       x = "Edition",
       y = "Number of Women Athletes")
df.women_participation %>%
  filter(Edition == 1900)
```


```{r}
# Animated bar chart race
writexl::write_xlsx(df.top10, "Top10.xlsx")
```



```{r}
# To excel for tableau
writexl::write_xlsx(df.summer, "TableauOlympics.xlsx")
writexl::write_xlsx(df.gold, "CountryMedals.xlsx")
writexl::write_xlsx(df.country, "TotalCountry.xlsx")
```

## Additional Analyses
*Any additional analysis should be conducted using Tableau.*

* * *

# Tableau Story

Embed your Tableau Story Here: *For the step-by-step guide on how to Embed, please go to our course site and click on the submission portal for the Data Visualization Project 2 assignment. In the assignment instructions, click on the link for the walk through (it is titled "Need help embedding your Tableau Dashboard in your R-Markdown file? Access the step-by-step instruction guide here on how to do this."). Be sure to Delete these instructions when completed.*



**USE THIS AREA FOR THE `<div> ... </div>` SECTION. DELETE THIS INSTRUCTION WHEN COMPLETED.**


**USE THIS CODE CHUNK FOR THE `<script> ... </script>` SECTION. DELETE THIS INSTRUCTION WHEN COMPLETED.**
```{js echo = FALSE}

```




URL Link to Tableau Story Site: [Your Name - Data Viz Project 2](**COPY AND PASTE YOUR TABLEAU STORY URL HERE**)

* * *

# References
*Include any references that you might have used to construct your graphic. This may include, but is not limited to citations for: graphics used, coding techniques, data collection, R-Markdown/RStudio documentation, etc. Delete this when complete.*

1. Reference 1
2. Reference 2
3. Reference 3
4. Reference 4
5. Reference 5

* * *

# All done!
Knit the completed R Markdown file as a HTML document (click the "Knit" button at the top of the script editor window) and upload it to the submission portal on Canvas.
